Model to study co-evolution of trade an culture

### 2015-03-29 article.tex
swap the two first state of the art sections and simplify the last one

@@

Within the archaeological field a large number of works have been proposed to quantify the properties of the economical interactions in the ancient Mediterranean societies~\cite{hopkins_taxes_1980,temin_market_2001,terpstra_trade_2011,minc_monitoring_2006,temin_economy_2006,wilson_approaches_2009,scheidel_model_2007,kessler_organization_2007}. In the works of~\cite{temin_market_2001,temin_economy_2006,wilson_approaches_2009}, based on the analysis of archaeological evidence, the Roman economy has been classified as a type of market economy where a variety of small tightly connected markets had loose connections between them. This view of a market system in ancient societies has been further expended in~\cite{minc_monitoring_2006} by a model applicable to various cases and studied in a test case of the aztec economical structure. The reaming works attempted at defining more exactly which type market economy it was, i.e. was it close to ours or remote ?

@@

The second component of our model is focused on the dynamics of cultural change. They comprise the collection of processes that promote or inhibit the spread of information by social interaction within a population \cite[3]{boyd_origin_2005}. This evolutionary framework is being used by an increasing number of social scientists, fostering the development of transdisciplinary efforts designed to understand cultural evolution \cite{henrich_evolution_2003}. Archaeology has a rich tradition of evolutionary studies \cite{lycett_cultural_2015}, exploring concepts such as gene-culture coevolution \cite{burger_absence_2007}, diffusion processes \cite{fort_synthesis_2012}, phylogenetic analysis of material culture \cite{obrien_cladistics_2001} or detection of rates of change in the archaeological record \cite{premo_cultural_2014}.

@@

\todo{speak about the fact that economy looks at convergence and culture looks before}

To the best of our knowledge two works have previously proposed frameworks that could be used to model the co-evolution of trade and culture. The model in~\cite{bentley_specialisation_2005} is proposed to study the wealth inequalities in a dynamic network. In this model the agents are in a virtual environment and can produce, consume and trade two products. This model has two major constraints which makes it unfit for the study presented here. First, the products are only for trade purposes and are not used as cultural factors, i.e. some goods can not be accumulated so that the agent show them to others. The cultural aspects appear only through the price associated to each product. Second, only two goods are considered and the system can not scale up to an undetermined number of products without major modifications.
% Agents can negotiate the rate at which the two products will be traded. The agents have a will (hard-coded in their behaviour) to keep their wealth and number of agents within their trading network above a threshold.

The model proposed in~\cite{macmillan_agent-based_2008} studies the settlement of primitive agricultural societies. In this model agents produce, trade and consume 2 essentials goods and 2 non-essentials goods. This model is interesting since it introduce the notion of non-essential goods which can be used to reflect the culture of a society. However, this framework is constrained as it does not allow the evolution of the trading network. Moreover the behaviour chosen for the agents is rather complex and not justified by observations made in economy or evolution of culture.

The model presented in this work proposes to integrate the most interesting features from~\cite{bentley_specialisation_2005} and~\cite{macmillan_agent-based_2008}.

\xrc{here a research question should be defined, and some hypotheses, e.g.: \emph{Did trade networks between regions bias cultural change towards anti-conformism?} or \emph{What is the relation between shipment time and cultural change?}}

\xrc{A co-evolutionary model is really difficult to study, so a careful justification needs to be done on why this is here needed}

>>>>

The first aspect of our model is focused on the dynamics of cultural change. They comprise the collection of processes that promote or inhibit the spread of information by social interaction within a population \cite[3]{boyd_origin_2005}. This evolutionary framework is being used by an increasing number of social scientists, fostering the development of transdisciplinary efforts designed to understand cultural evolution \cite{henrich_evolution_2003}. Archaeology has a rich tradition of evolutionary studies \cite{lycett_cultural_2015}, exploring concepts such as gene-culture coevolution \cite{burger_absence_2007}, diffusion processes \cite{fort_synthesis_2012}, phylogenetic analysis of material culture \cite{obrien_cladistics_2001} or detection of rates of change in the archaeological record \cite{premo_cultural_2014}.

@@

The second aspect of our model is focused on the dynamics produced by trade in ancient Mediterranean societes. Within the archaeological field a large number of works have been proposed to quantify the properties of the economical interactions in the ancient Mediterranean societies~\cite{hopkins_taxes_1980,temin_market_2001,terpstra_trade_2011,minc_monitoring_2006,temin_economy_2006,wilson_approaches_2009,scheidel_model_2007,kessler_organization_2007}. In the works of~\cite{temin_market_2001,temin_economy_2006,wilson_approaches_2009}, based on the analysis of archaeological evidence, the Roman economy has been classified as a type of market economy where a variety of small tightly connected markets had loose connections between them. This view of a market system in ancient societies has been further expended in~\cite{minc_monitoring_2006} by a model applicable to various cases and studied in a test case of the aztec economical structure. The reaming works attempted at defining more exactly which type market economy it was, i.e. was it close to ours or remote ?

@@

@@

\jm{tried a justification of co-evolutionary framework but I find it too light. ideas ?}

Ultimately we aim to study the relation between cultural change and trade networks within the Roman Empire. Examples of questions that should be answered by our model are: What is the relation between distances in the trade network and cultural changes ? Did anti-conformism promote the adaptation of new products coming from remote provinces ? In order to answer these questions the model used has to take into account both cultural effects and trade influence. More precisely, the variations in these two aspects should be modelled and should influence each other.  Therefore, in order to answer the questions targeted we need a co-evolutionary model able to take into account both trade and cultural changes.

To the best of our knowledge two works have previously proposed frameworks that could be used to model the co-evolution of trade and culture~\cite{bentley_specialisation_2005,macmillan_agent-based_2008}. The first proposes a relatively simple model which is limited by 2 products, a fixed behaviour of the agents and a limited variation of cultural mechanisms. The second proposes a more complete model including a large number of goods (certain being essential and the others non-essential), but suffer from a greater complexity and the absence of the notion of trade networks. Moreover, both models propose new dynamics but these dynamics are not compared to the state of the art on the models for cultural change and trade.

The model presented in this work aims to offer a framework for the co-evolution of trades and culture which relies on a low number of parameters. In order to assess its validity we will reproduce results obtained in the literature and verify that the results obtained are conform to what is expected.
